---
title: "rangewide_excel_datasheet_ingest"
author: "Patrick D. lorch"
date: "2023-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Todo

* fix survey conversion (see comment around line 154)
* Get rid of blank rows
* Get to run on dir with multiple files and check

## Libraries


```{r libs}
library(readxl)
library(openxlsx)
library(dplyr)
library(tidyr)
library(purrr)
```

## Input variables

You want to run one of the path setting variables, then run the file.list and path.list steps below for each directory.

You may need to replace "\" with "/" in the path name if you use Copy as path on windows to locate directory with the .xlsx files in them for each Other entity.

```{r variables}
# USFS
rangewide_files_path = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/Other Entity Data/NM Other Entity/USFS"

# NM BdA NWR
rangewide_files_path = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/Other Entity Data/NM Other Entity/BdA NWR/Survey Forms"

# BLM - Cody Howard
rangewide_files_path = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/Other Entity Data/NM Other Entity/BLM - Cody Howard"

# Ecosphere
rangewide_files_path = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/Other Entity Data/NM Other Entity/Ecosphere"

# Matt Johnson - TNC
rangewide_files_path = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/Other Entity Data/NM Other Entity/Matt Johnson - TNC"

# Sevilleta NWR
rangewide_files_path = "C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/YBCU - Rangewide Surveys CSWG/Other Entity Data/NM Other Entity/Sevilleta NWR"


file.list = list.files(path = rangewide_files_path, pattern='*.xlsx')
path.list = file.path(rangewide_files_path, file.list)
```

## Functions

These should only need to be sourced once.

```{r functions}
# Headers
read_header = function(path = path.list, 
                       sheet = sheet0,
                       range = headercells,
                       col_names = F){
  read_excel(path, sheet, range, col_names)
}

# This maps the data into the right column and renames
header_extracts = function(header = headers.df.list){
  headers.df = data.frame(SiteName = as.character(header[2,3]),
                        County = as.character(header[2,11]),
                        Elevation_m = as.numeric(header[2,18]),
                        State = as.character(header[2,22]),
                        USGS_QuadName = as.character(header[3,3]),
                        CrkRvrWtlndCnynLkName = as.character(header[3,16]),
                        UTMzone = as.character(header[5,9]),
                        StartUTM_Easting = as.numeric(header[5,15]),
                        StartUTM_Northing = as.numeric(header[5,18]),
                        MagneticDeclinationDeg = as.numeric(header[6,23]),
                        Datum = as.character(header[6,9]),
                        StopUTM_Easting = as.numeric(header[6,15]),
                        StopUTM_Northing = as.numeric(header[6,18]),
                        SiteSurveyedPrev = as.character(header[7,12]),
                        PreviousYearName = as.character(header[7,18])
                     )
}

# Survey and Survey point data
read_survpts = function(path = path.list, 
                       sheet = sheet0,
                       range = surveycells,
                       col_names = F){
  read_excel(path, sheet, range, col_names)
}

# Surveys
survey_extracts = function(surveys = survpts.df.list) {
  surveys = surveys[, 1:6]
  start = 4
  skip = 8
  df.list = list()
  surveys.list = for (i in 0:4) {
    ind = start + i * skip
    df.list[[i + 1]] = data.frame(
      SurveyNumber = as.numeric(surveys[ind, 1]),
      SurveyDate = convertToDateTime(surveys[ind, 4]),
      Observers = paste(surveys[(ind + 2), 1],
                        surveys[(ind + 3), 1],
                        surveys[(ind + 4), 1],
                        surveys[(ind + 5), 1],
                        surveys[(ind + 6), 1], 
                        sep = "|"),
      SurveyStart = convertToDateTime(as.numeric(surveys[ind, 4]) +
                                        as.numeric(surveys[ind + 2, 4])),
      SurveyStop = convertToDateTime(as.numeric(surveys[ind, 4]) +
                                       as.numeric(surveys[ind + 4, 4]))
    )
    return(df.list)
   }
  return(data.frame(surveys.list))

}

# Survey Points
survpts_extracts = function(survpts = survpts.df.list){
  survnames = c("SurveyNumber", "YBCU_No", "DetectionTime", "DetectionMethod",
                "DetectionType", "VocalizationType", "NumberOfPlaybacks",
                "BehaviorObserved", "UTM_E", NA, "UTM_N", NA, "Distance_m",
                "Bearing", "CorrectedUTM_E", "CorrectedUTM_N", NA,
                "TrigBearing", "E_correction", "N_correction",
                "CorrectedUTM_E_2", "CorrectedUTM_N_2", "State", "Sitename")
  survpts = survpts[-c(1,2), 7:30]

  survpts.df = data.frame(survpts)
  names(survpts.df) = survnames
  
  return(survpts.df)
}

```


## Header import

Good options
https://www.r-bloggers.com/2021/06/reading-data-from-excel-files-xlsxlsxcsv-into-r-quick-guide/

Make sure you have the sheet closed before running this.

```{r header}
headercells = "A1:W7"
sheet0 = "BlanKForm"

# Get list of spreadsheets
headers.df.list = lapply(path.list, read_header)
# extract data we need and output list of 1 line dataframes
headerdata.list = lapply(headers.df.list, header_extracts)
# Combine rows from list
headerdata.df = do.call(rbind, headerdata.list)
write.csv(headerdata.df, file.path(rangewide_files_path, "headerdata.csv"))
```

## Surveys

```{r surveys}
surveycells = "A9:AB50"
# Make sure to run the header part above to reset this
surveystates = headerdata.df$State
surveysitenames = headerdata.df$SiteName
sheet0 = "BlanKForm"

# Get list of data.frames
# extract data we need and output list of 1 line dataframes
survdata.list = lapply(survpts.df.list, survey_extracts)
survdata.list = purrr::list_flatten(survdata.list)
survdata.list = mapply(cbind, survdata.list, 
                       State = surveystates,
                       SiteName = surveysitenames, 
                       SIMPLIFY = F)
# Combine rows from list
survdata.df = do.call(rbind, survdata.list)
write.csv(survdata.df, file.path(rangewide_files_path, "survdata.csv"))

```


## Survey Points import


```{r surveypoints}
surveycells = "A9:AB50"
# Make sure to run the header part above to reset this
surveystates = headerdata.df$State
surveysitenames = headerdata.df$SiteName
sheet0 = "BlanKForm"

# Get list of data.frames
# extract data we need and output list of 1 line dataframes
survpts.df.list = lapply(path.list, read_survpts)
survpts.df.list = mapply(cbind, survpts.df.list,
                         State = surveystates,
                         SiteName = surveysitenames,
                         SIMPLIFY = F)
survptdata.list = lapply(survpts.df.list, survpts_extracts)
# Combine rows from list
survptdata.df = do.call(rbind, survptdata.list)
write.csv(survptdata.df, file.path(rangewide_files_path, "survptdata.csv"))

```

